### 2010-06-22 trunk/paper/IIWAS2010/content/system_arch.tex
Modifications to category divergence section.

To answer a regular user's queries that is in SSQ format (a \textit{candidate
SSQ}), Morpheus finds similar SSQs that belong to QRMs in the store (a
\textit{qualified SSQ}). We need a similarity measure to match the candidate SSQ with a qualified SSQ within the Morpheus data store. For the SSQ similarity, we consider the SSQ components [\ref{sec:morpheus}] such as query-realm, input terms and output terms, and their assigned categories or classes. Thus, Class Divergence, a similarity measure, is defined for these classes based on ontology's class heterarchy. It is motivated by the concept of multiple dispatch in CLOS \cite{Steele1990} and Dylan programming \cite{Barrett1996} for generic function type matches. Once we have the similarity measures between the candidate SSQ and qualified SSQs in the QRM data store, we order the relevant SSQs and associated QRMs. The order provides a ranking for the results to the user. Subsections \ref{sec:ctd}, and \ref{sec:oqom} describe this in detail.

\subsubsection{Class Divergence}
\label{sec:ctd}

We employ a measure of compatibility, which we call \textit{class divergence},
between a source class and a target class using the \textit{topological
structure} of the classes in an ontology. We write $S \prec T$ for the reflexive
transitive closure of the superclass relation. Let $d(P,Q)$ represent the hop
distance in the directed ontology inheritance graph from $P$ to $Q$. The
divergence between a source and target class ranges from zero (for identical
classes) to one (for incompatible classes). Let $S$ be the source class, $T$ be
the target class, and let $C$ be a common ancestor class of $S$ and $T$
minimizing $d(S,C) + d(T,C)$. The class divergence between $S$ and $T$, the
unnormalized class dissimilarity, is defined given by:

\begin{equation}
d_{class}(S, T) = \begin{cases}
0 & S.{Uri} \equiv T.{Uri}\\
d(S, T)/(3h) & S \prec T\\
1 & T \prec S\\
(d(S,root) + d(S,C) \\ \ \ \ \ + d(T,C))/(3h) & otherwise
\end{cases}
\end{equation}

@@

Note, if $S \prec T$ and $S \not\prec Q$ then $d_{class}(S,T) <
d_{class}(S,Q)$, that is, the divergence of a source class to a target
ancestor class is smaller than the divergence of a source class to any
class that is not an ancestor. This is an important property in
determining the compatibility of classes for answering queries.  If a
SSQ answers queries concerning an ancestor class, it is more relevant
that a SSQ that answers queries from any non-ancestral class.

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=80mm]{img/class_divergence.eps}
\caption{\textbf{An abbreviated graphical representation of an automotive
ontology}: This is a manually created ontology, rooted at the class Ontology
Root, from the Wikipedia category hierarchy and live Wikipedia pages. The arrows
represent the inheritance relationships between the classes in a topology.}
\label{fig:class_divergence}
\end{figure}

Figure \ref{fig:class_divergence} illustrates the calculation of class
divergence. Suppose we want to find the class divergence
$d_{class}$\textit{(Automotive
  Industry, Racing Cars)}. The \textit{Automotive Industry} is the
subclass of \textit{Automobiles} and \textit{Racing Cars} is the
subclass of \textit{Auto Racing}, which is a subclass of \textit{Automobiles}. 
Therefore, the common ancestor for
these two classes is \textit{Automobiles} ($C$). They share
the path from \textit{Automobiles} to the ontology root. So, the unnormalized
divergence \textit{d(Automotive Industry, Ontology Root)} is 32 that is
calculated from the path to the \textit{Ontology Root} class,
\textit{d(Automotive Industry,
Automobiles)} is one
(dashed arrow), and \textit{d(Racing Cars, Automobiles)} is two
(dashed arrow). Suppose, the calculated ontology tree height is 294. Thus,
$d_{class}$\textit{(Automotive Industry, Racing Cars)} is 35, 
which corresponds to a normalized value of $35/3*294$.

>>>>

To answer a regular user's queries that is in SSQ format (a \textit{candidate
SSQ}), Morpheus finds similar SSQs that belong to QRMs in the store (a
\textit{qualified SSQ}). We need a similarity measure to match the candidate SSQ
with a qualified SSQ within the Morpheus data store. For the SSQ similarity, we
consider the SSQ components [\ref{sec:morpheus}] such as query-realm, input
terms and output terms, and their assigned categories or classes. Thus,
\textit{category divergence}, a similarity measure, is defined for these classes
based on ontology's class heterarchy. It is motivated by the concept of multiple
dispatch in CLOS \cite{Steele1990} and Dylan programming \cite{Barrett1996} for
generic function type matches. Once we have the similarity measures between the
candidate SSQ and qualified SSQs in the QRM data store, we order the relevant
SSQs and associated QRMs. The order provides a ranking for the results to the
user. Subsections \ref{sec:ctd}, and \ref{sec:oqom} describe this in detail.

\subsubsection{Catgeory Divergence}
\label{sec:ctd}

We employ a measure of compatibility, which we call \textit{category
divergence},
between a source category and a target category using the \textit{topological
structure} of the categories in an ontology. We write $S \prec T$ for the
reflexive
transitive closure of the supercategory relation. Let $d(P,Q)$ represent the hop
distance in the directed ontology inheritance graph from $P$ to $Q$. The
divergence between a source and target category ranges from zero (for identical
categories) to one (for incompatible categories). Let $S$ be the source
category, $T$ be
the target category, and let $C$ be a least common ancestor category of $S$ and
$T$
minimizing $d(S,C) + d(T,C)$. The category divergence between $S$ and $T$ is
defined given by:

\begin{equation}
d_{cat}(S, T) = \begin{cases}
0 & S.{Uri} \equiv T.{Uri}\\
d(S, T)/(3h) & S \prec T\\
1 & T \prec S\\
(d(S,root) + d(S,C) \\ \ \ \ \ + d(T,C))/(3h) & otherwise
\end{cases}
\end{equation}

@@

Note, if $S \prec T$ and $S \not\prec Q$ then $d_{cat}(S,T) <
d_{cat}(S,Q)$, that is, the divergence of a source category to a target
ancestor category is smaller than the divergence of a source category to any
category that is not an ancestor. This is an important property in
determining the compatibility of category for answering queries.  If a
SSQ answers queries concerning an ancestor category, it is more relevant
that a SSQ that answers queries from any non-ancestral category.

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=90mm]{img/automotive_ontology.eps}
\caption{An abbreviated graphical representation of an automotive
ontology: This is a manually created ontology, from the Wikipedia category
hierarchy and pages. The edges represent inheritance relationships between the
category in a topology.}
\label{fig:automotive_ontology}
\end{figure}

Suppose we want to find the category divergence between \textit{Toyota Camry}
and \textit{Coupes} [Figure \ref{fig:automotive_ontology}]. 
\textit{Toyota Camry} is the subcategory of \textit{Sedans}, and \textit{Coupes}
is the subcategory of \textit{Cars}. Therefore, the least common ancestor ($C$)
for these two categories is \textit{Cars}. In addition, the tree height is $5$.
So, the normalized divergence $d_{cat}$\textit{(Toyota Camry, Coupes)} is $7/15$
that is calculated from \textit{d(Toyota Camry, Root)} is 5, \textit{d(Toyota
Camry, Cars)} is two, and \textit{d(Coupes, Cars)} is one.
